An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-31 VLDB2026/main.tex
Update main.tex
1

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
%				\textit{AND} and \textit{OR} indicate whether the algorithm supports the corresponding logical operations.
%				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
%				\textit{Complex Boolean} indicates if the algorithm supports complex expressions, such as using the NOT operator or nested logic with parentheses.
%				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
%				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
%				\textit{Disk Support} indicates whether the algorithm supports disk storage.
%				\textit{Base} indicates the foundational ANN algorithm for the hybrid query method.
		\textcolor{orange}{
		\textit{Filtered}: FilteredVamana, \textit{Stitched}: StitchedVamana, \textit{Win}: WinFilter.
		\textit{AND} / \textit{OR}: Whether the algorithm supports the corresponding logical operations.
		\textit{Flexible Attributes}: Allows the number of query attributes to differ from that in index construction.
		\textit{Complex Boolean}: Supports NOT or nested Boolean expressions.
		\textit{Dynamic Insert}: Supports dynamic data insertion.
		\textit{Multi Thread}: Supports multi-threaded search.
		\textit{Disk Support}: Supports disk-based storage.
		\textit{Base}: Underlying ANN index structure used.}
		\textit{A}: Post-filtering, 
		\textit{B}: Pre-filtering, 
		\textit{C}: Simultaneous filtering, 
		\textit{Y}: Support, 
		\textit{N}: Unsupport. 
		\end{minipage}}

@@

Faiss\footnote{\textcolor{blue}{https://github.com/facebookresearch/faiss}} \cite{Faiss} is an efficient similarity search library. It uses an ID selector to filter data before performing approximate nearest neighbor search. This supports hybrid search.In Faiss, we use the HQI batch optimization strategy in the IVF index. It groups queries with the same filter conditions. Each group only runs the filter once. It uses efficient matrix operations to compute vector similarity in batches.

PASE\footnote{\textcolor{blue}{https://github.com/alipay/PASE}} \cite{pase} is a vector index plugin for PostgreSQL. It uses a "search first, then filter" strategy. But because the candidate set size is fixed, it may not return enough top-k results.

VBASE\footnote{\textcolor{blue}{https://github.com/microsoft/MSVBASE}} \cite{vbase} is also a vector plugin for PostgreSQL. It applies filters during the HNSW index traversal. This ensures enough results.

Milvus\footnote{\textcolor{blue}{https://github.com/milvus-io/milvus}} \cite{milvus} is a vector database built for large-scale similarity search. It filters data first, then performs approximate nearest neighbor search. This supports hybrid search.

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.}
	% Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, the range filtering experiment only uses one attribute, and the dataset is sorted in ascending order according to the attribute value size. WIT-Real and YT-Audio-Real use real attribute values, and the rest use randomly generated attribute values.}
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.} 
	Based on this definition, we evaluate algorithm performance under four query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. For space considerations, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. 
%	The experimental results show consistent performance across all datasets and query ranges.

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}} 
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:range-cross-platform}. Although the third-ranked algorithm differs, we consider the ranking order remains stable since the performance of the third and fourth algorithms is comparable. So our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig: range big dataset}, On the deep100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

@@

\begin{table*}[htbp]
	\caption{Algorithm Recommendation}
	\resizebox{\textwidth}{!}{
\begin{tabular}{|l|*{16}{c|}}
	\hline
	\multirow{2}{*}{\textbf{Application Scenarios / Algorithms}} 
	& \multicolumn{6}{c|}{\textbf{AF}} 
	& \multicolumn{5}{c|}{\textbf{RF}} 
	& \multicolumn{5}{c|}{\textbf{AF / RF}} \\
	\cline{2-17}
	& \textbf{NHQ} & \textbf{Filtered} & \textbf{Stitched} & \textbf{CAPS}  & \textbf{UNG} & \textbf{Puck} 
	& \textbf{DSG} & \textbf{iRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{Win} 
	& \textbf{ACORN}& \textbf{Faiss} & \textbf{Milvus} & \textbf{VBASE} & \textbf{PASE} \\
	\hline
	\textbf{Query Performance} & S & M & S & M &\textbf{S} & S & M & M & W & M & S & M/M & M/M & M/M & W/W & W/W \\
	\textbf{Index Time} &M  &M  &M  &M  &M  &S  &W  &M  &S  &W  &W  &W/W &\textbf{S}/\textbf{S} &S/S  &W/M  &W/W  \\
	\textbf{Peak Memory} &W  &M  &M  &M  &M  &S  &W  &W  &M  &W  &W  &W/W  &\textbf{S}/\textbf{S}  &S/M  &S/M  &S/S \\
	\textbf{Large-Scale Dataset} & S & S & W & W &\textbf{S} &S  &M  &S  &M  &S  &\textbf{S}  &M/W  &W/M  &W/W  &M/M  &-  \\
	\textbf{Rare Attribute \& Low Range} &M  &M  &S  &S &\textbf{S}  &S  &S  &S  &W  &M  &\textbf{S} &M/M  &M/M  &W/W  &W/W  &W/W  \\
	\textbf{Robust to Multi-Modal Dataset} &S  &S  &W  &S  &\textbf{S}  &M  &M  &S  &M  &S  &\textbf{S}  &M/M  &W/M  &W/W  &W/W  &W/W  \\
	\textbf{Robust to Attribute Distribution} &W  &W  &S  &S  &S  &S  &-  &-  &-  &-  &-  &S/-  &W/-  &\textbf{S}/-  &W/-  &W/-  \\
	\textbf{High LID Dataset} &S  &M  &S  &M  &\textbf{S}  &M  &M  &S  &M  &S  &\textbf{S}  &W/W  &M/S  &W/W  &W/M  &W/W  \\
	\textbf{General-Purpose} &\faCheck  &\faCheck  &\faTimes  &\faTimes  &\faCheck  &\faTimes  &\faTimes  &\faTimes &\faTimes  & \faCheck &\faCheck&\faTimes  &\faCheck  &\faCheck  &\faCheck&\faTimes  \\
	\hline
\end{tabular}
	}
	\vspace{1mm}
\begin{flushleft}
	\textbf{Performance Levels:} \quad \textbf{S} = Strongest, S = Strong, M = Medium, W = Weak, - = Not Supported \\
	\textbf{Recommendation:} \quad \faCheck = Recommended, \faTimes = Not Recommended
\end{flushleft}
	\centering

>>>>

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
%				\textit{AND} and \textit{OR} indicate whether the algorithm supports the corresponding logical operations.
%				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
%				\textit{Complex Boolean} indicates if the algorithm supports complex expressions, such as using the NOT operator or nested logic with parentheses.
%				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
%				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
%				\textit{Disk Support} indicates whether the algorithm supports disk storage.
%				\textit{Base} indicates the foundational ANN algorithm for the hybrid query method.
		\textcolor{orange}{
		\textit{Filtered}: FilteredVamana, \textit{Stitched}: StitchedVamana, \textit{Win}: WinFilter.
		\textit{AND} / \textit{OR}: Whether the algorithm supports the corresponding logical operations.
		\textit{Flexible Attributes}: Allows the number of query attributes to differ from that in index construction.
		\textit{Complex Boolean}: Supports NOT or nested Boolean expressions.
		\textit{Dynamic Insert}: Supports dynamic data insertion.
		\textit{Multi Thread}: Supports multi-threaded search.
		\textit{Disk Support}: Supports disk-based storage.
		\textit{Base}: Underlying ANN index structure used.}
		\textit{A}: Post-filtering (Filter after ANN search), 
		\textit{B}: Pre-filtering (Filter before ANN search), 
		\textit{C}: Simultaneous filtering (Filter simultaneously with ANN search), 
		\textit{Y}: Support, 
		\textit{N}: Unsupport. 
		\end{minipage}}

@@

\noindent\textbf{\underline{Faiss}\footnote{\textcolor{blue}{https://github.com/facebookresearch/faiss}} \cite{Faiss}}
Faiss is an efficient similarity search library. It uses an ID selector to filter data before performing ANN search. This supports hybrid search. In Faiss, we use the HQI batch optimization strategy in the IVF index. It groups queries with the same filter conditions. Each group only runs the filter once. It uses efficient matrix operations to compute vector similarity in batches.

\noindent\textbf{\underline{PASE} \footnote{\textcolor{blue}{https://github.com/alipay/PASE}} \cite{pase}} PASE is a vector index plugin for PostgreSQL. It uses a post-filtering strategy. Since the candidate set size is fixed, it may fail to return enough top-k results.

\noindent\textbf{\underline{VBASE} \footnote{\textcolor{blue}{https://github.com/microsoft/MSVBASE}} \cite{vbase}} VBASE is also a vector plugin for PostgreSQL. It applies filters during the HNSW index traversal. This ensures enough results.

\noindent\textbf{\underline{Milvus} \footnote{\textcolor{blue}{https://github.com/milvus-io/milvus}} \cite{milvus}} Milvus is a vector database designed for large-scale similarity search. In hybrid queries, it adopts a pre-filtering strategy, similar to the ID selector mechanism used in Faiss.
%It filters data first, then performs approximate nearest neighbor search. This supports hybrid search.

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.}
	% Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{blue}{Due to the limitations of existing range filtering algorithms, the range filtering experiment only uses one attribute, and the dataset is sorted in ascending order according to the attribute value size. WIT-Real and YT-Audio-Real use real attribute values, and the rest use randomly generated attribute values.}
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.} 
	Based on this definition, we evaluate algorithm performance under four query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. For space considerations, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. The experimental results show consistent performance across all datasets and query ranges.

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}} 
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:range-cross-platform}. Although the third-ranked algorithm differs, we consider the ranking order essentially unchanged since the performance of the third and fourth algorithms is comparable. So our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig: range big dataset}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

@@

\begin{table*}[htbp]
	\caption{Algorithm Recommendation}
	\resizebox{\textwidth}{!}{
\begin{tabular}{|l|*{16}{c|}}
	\hline
	\multirow{2}{*}{\textbf{Application Scenarios / Algorithms}} 
	& \multicolumn{6}{c|}{\textbf{AF}} 
	& \multicolumn{5}{c|}{\textbf{RF}} 
	& \multicolumn{5}{c|}{\textbf{AF / RF}} \\
	\cline{2-17}
	& \textbf{NHQ} & \textbf{Filtered} & \textbf{Stitched} & \textbf{CAPS}  & \textbf{UNG} & \textbf{Puck} 
	& \textbf{DSG} & \textbf{iRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{Win} 
	& \textbf{ACORN}& \textbf{Faiss} & \textbf{Milvus} & \textbf{VBASE} & \textbf{PASE} \\
	\hline
	\textbf{Query Performance} & S & M & S & M &\textbf{S} & S & M & M & W & M & S & M/M & M/M & M/M & W/W & W/W \\
	\textbf{Index Time} &M  &M  &M  &M  &M  &S  &W  &M  &S  &W  &W  &W/W &\textbf{S}/\textbf{S} &S/S  &W/M  &W/W  \\
	\textbf{Peak Memory} &W  &M  &M  &M  &M  &S  &W  &W  &M  &W  &W  &W/W  &\textbf{S}/\textbf{S}  &S/M  &S/M  &S/S \\
	\textbf{Large-Scale Dataset} & S & S & W & W &\textbf{S} &S  &M  &S  &M  &S  &\textbf{S}  &M/W  &W/M  &W/W  &M/M  &-  \\
	\textbf{Rare Attribute \& Low Range} &M  &M  &S  &S &\textbf{S}  &S  &S  &S  &W  &M  &\textbf{S} &M/M  &M/M  &W/W  &W/W  &W/W  \\
	\textbf{Robust to Multi-Modal Dataset} &S  &S  &W  &S  &\textbf{S}  &M  &M  &S  &M  &S  &\textbf{S}  &M/M  &W/M  &W/W  &W/W  &W/W  \\
%	\textbf{Robust to Attribute Distribution} &W  &W  &S  &S  &S  &S  &-  &-  &-  &-  &-  &S/-  &W/-  &\textbf{S}/-  &W/-  &W/-  \\
	\textbf{High LID Dataset} &S  &M  &S  &M  &\textbf{S}  &M  &M  &S  &M  &S  &\textbf{S}  &W/W  &M/S  &W/W  &W/M  &W/W  \\
	\textbf{General-Purpose} &\faCheck  &\faCheck  &\faTimes  &\faTimes  &\faCheck  &\faTimes  &\faTimes  &\faTimes &\faTimes  & \faCheck &\faCheck&\faTimes  &\faCheck  &\faCheck  &\faCheck&\faTimes  \\
	\hline
\end{tabular}
	}
	\vspace{1mm}
\begin{flushleft}
	\textbf{Performance Levels:} \quad \textbf{S} = Strongest, S = Strong, M = Medium, W = Weak, - = Not Supported \\
	\textbf{Recommendation:} \quad \faCheck = Recommended, \faTimes = Not Recommended
\end{flushleft}
	\centering
